### 2022-12-09 paper1.tex
one-sided sceptical p

\section*{Abstract}
Replication studies are increasingly conducted in order to confirm original
findings. However, there is no established standard how to assess replication
success and in practice many different approaches are used. The purpose of this
paper is to refine and extend a recently proposed reverse-Bayes approach for the
analysis of replication studies. We show how this method is directly related to
the relative effect size, the ratio of the replication to the original effect
estimate. This perspective leads to a new proposal to recalibrate the assessment
of replication success, the golden level. The recalibration ensures that for
borderline significant original studies replication success can only be achieved
if the replication effect estimate is larger than the original one. Conditional
power for replication success can then take any desired value if the original
study is significant and the replication sample size is large enough. Compared
to the standard approach to require statistical significance of both the
original and replication study, replication success at the golden level offers
uniform gains in project power and controls the Type-I error rate if the
replication sample size is not smaller than the original one. An application to
data from four large replication projects shows that the new approach leads to
more appropriate inferences, as it penalizes shrinkage of the replication
estimate compared to the original one, while ensuring that both effect estimates
are sufficiently convincing on their own.

\textbf{Key words}: Power, replication studies, sceptical $p$-value, shrinkage,
two-trials rule, type-I error rate

@@

In this paper we first review the \citet{Held2020} approach for the assessment
of replication success, followed by showing how it relates to the relative
effect size (Section \ref{sec1:res}). This perspective is used in Section
\ref{sec1:goldenthresh} and \ref{sec1:recalib} to propose a recalibration of the
method, the \textit{golden level}, which leads to a more appropriate criterion
for replication success compared to the two-trials rule (Section \ref{sec1:2TR}).
In Section \ref{sec1:ER} we study power and Type-I error rates of the proposed
method and compare it to the two-trials rule. The recalibrated method ensures
that conditional power can take any desired value if the original study has been
significant and the replication sample size is large enough (Section
\ref{sec1:powerrep}), controls the overall Type-I error if the replication sample
size is not smaller than the original one (Section \ref{sec1:T1E}), and offers
uniform gains in project power compared to the two-trials rule (Section
\ref{sec1:PP}). Section \ref{sec1:application} describes an application to data
from four replication projects and Section \ref{sec1:discussion} closes with some
discussion.

@@

\section{Power and Type-I Error Rate}\label{sec1:ER}
Although Bayesian methods do not rely on the frequentist paradigm of repeated
testing, it is still useful to investigate their frequentist operating
characteristics \citep{Dawid1982, Rubin1984, Grieve2016} and this also holds for
the proposed reverse-Bayes assessment of replication success. We first condition
on the results from the original study and compare the power to achieve
replication success with the two-trials rule in Section \ref{sec1:powerrep}. We
then assume that none of the two studies have been conducted and investigate the
overall Type-I error rate (Section \ref{sec1:T1E}) and the project power (Section
\ref{sec1:PP}) \citep{Maca2002} over both studies in combination for fixed
relative sample size $c$.

@@

\subsection{Overall Type-I error rate}\label{sec1:T1E}
The two studies are assumed to be independent with Type-I error rate fixed at
$\alpha$ for each of them, so the Type-I error rate of the two-trials rule over
the entire project is simply $\alpha^2$ for any value of the relative effect
size $c$. In contrast, the Type-I error rate of the proposed replication success
assessment {depends on} the relative sample size $c$.

For $c=1$, \citet[Section 3]{Held2020} showed that $z_S^2$ in
\eqref{eq1:extrinsic.p2} simplifies to half the harmonic mean of the squared test
statistics $z_o^2$ and $z_r^2$. The connection $z_S^2 = z_H^2/4$ to the harmonic
mean $\chi^2$-test statistic $z_H^2$ \citep{Held2020b}, which has a
$\chi^2(1)$-distribution under the null hypothesis, makes it straightforward to
compute the Type-I error rate at level $\alpha_S$ for $c = 1$ as
\begin{eqnarray}\label{eq1:T1E}
\mbox{T1E} =   \left\{1-\Phi\left[2 \, \Phi^{-1}\left(1-\alpha_S \right)
 \right]\right\}/2.
\end{eqnarray}
For the golden level $\alpha_S =0.062$ at $\alpha = 0.025$, the Type-I error
rate \eqref{eq1:T1E} is $0.0515$\%, slightly less than the Type-I error rate
$\alpha^2=0.0625$\% of the two-trials rule. For comparison, the Type-I error
rate at the nominal level $\alpha_S=0.025$ is $0.0022$\%, much smaller than
0.0625\%.

For $c \neq 1$, the Type-I error rate can be calculated through numerical
integration:
\begin{equation}\label{eq1:T1E0}
  \mbox{T1E} = \int_{z_{\alpha_S}}^{\infty}
\P(z_r \geq \zrmin \given z_o, c, \alpha_S) \,
  \phi(z_o) \, dz_o,
\end{equation}
where $\phi(\cdot)$ denotes the standard normal density function. The first term
in the integral of \eqref{eq1:T1E0} is the probability of replication success at
level $\alpha_S$ conditional on a fixed original test statistic $z_o$ and a
relative sample size $c$. Now $z_r \sim \Nor(0,1)$ under the null hypothesis, so
this term simplifies to
$\P(z_r \geq \zrmin \given z_o, c, \alpha_S) = 1 - \Phi(\zrmin)$ where $\zrmin$
in \eqref{eq1:eq.z} depends on $z_o$, $c$, and $\alpha_S$.

@@

\caption{Overall Type-I error rate (left) and project power (right) for fixed
  relative sample size $c$. Results are given for replication success at the
  nominal and golden level and compared with the two-trials rule (2TR) at
  $\alpha = 0.025$. The dashed darkgrey line is the project power at the golden
  level based on significant original studies ($p_o \leq 0.025$). The power of
  the original study is 90\%.}
\label{fig1:fig4}
\end{figure}

The left plot in Figure \ref{fig1:fig4} displays the Type-I error rate for
$\alpha=0.025$ as a function of the relative sample size $c$. It can be seen
that the Type-I error of the replication success approach decreases with
increasing relative sample size $c$. This also follows from \eqref{eq1:T1E0}
where $\P(z_r \geq \zrmin \given z_o, c, \alpha_S) = 1 - \Phi(\zrmin)$ decreases
with increasing $c$, because $\zrmin$ increases with increasing $c$, see
equation \eqref{eq1:eq.z}.

The Type-I error rate of the nominal level is always below the target
$0.0625$\%. Although the Type-I error will eventually attain $\alpha^2$ in the
limit $c \downarrow 0$ \citep[Section 3.4]{Held2020}, the nominal level seems to
be too stringent for realistic values of $c$. The Type-I error rate of the
golden level is smaller than $0.0625$\% for $c >0.85$. Appropriate Type-I error
control is thus ensured even for replication studies where the sample size is
slightly smaller than in the original study.

@@

\end{center}
\caption{Overall Type-I error rate if the replication sample size equal to the
  original study ($c=1$). The two-trials rule (2TR) is compared to replication
  success at the golden and nominal level for different values of $\alpha$.}
\label{fig1:fig5}
\end{figure}

Figure \ref{fig1:fig5} compares for $c=1$ the Type-I error rate \eqref{eq1:T1E} of
replication success at the golden and at the nominal level with the two-trials
rule for different values of $\alpha$. The Type-I error rate of the two-trials
rule is $\alpha^2$ and the replication success approach at the nominal level
always has a much smaller Type-I error rate than $\alpha^2$. At the golden level
the Type-I error rate of the replication success approach is much closer to
$\alpha^2$, still slightly smaller if $\alpha < 0.058$. For $\alpha = 0.058$ the
Type-I error rate is equal to the Type-I error rate $0.058^2 = 0.34\%$ of the
two-trials rule and for $\alpha > 0.058$ the Type-I error rate is slightly
larger than $\alpha^2$. The Type-I error rate for replication success decreases
with increasing $c$, so as long as the replication sample size is not smaller
than the original sample size, Type-I error control at $\alpha^2$ is guaranteed
at the golden level for any one-sided level $\alpha < 0.058$.

@@

We further refined the method by proposing the golden level, a new threshold for
replication success. It guarantees that borderline significant original studies
can only be replicated successfully if the replication effect estimate is larger
than the original one. Compared to the two-trials rule, the golden level offers
uniform gains in project power and controls the Type-I error rate at any
one-sided level $\alpha < 0.058$ if the replication sample size
is not smaller than the original one. Empirical evaluation of data from four
replication projects highlights that in most cases the methods are in agreement,
however, for the study pairs where the approaches disagree, the replication
success approach seems to lead to more sensible conclusions. The good
performance has been recently confirmed by a comparison of different replication
success metrics through a simulation study in the presence of publication bias
\citep{Muradchanian2021}.

>>>>

\section*{Abstract}
Replication studies are increasingly conducted in order to confirm original
findings. However, there is no established standard how to assess replication
success and in practice many different approaches are used. The purpose of this
paper is to refine and extend a recently proposed reverse-Bayes approach for the
analysis of replication studies. We show how this method is directly related to
the relative effect size, the ratio of the replication to the original effect
estimate. This perspective leads to a new proposal to recalibrate the assessment
of replication success, the golden level. The recalibration ensures that for
borderline significant original studies replication success can only be achieved
if the replication effect estimate is larger than the original one. Conditional
power for replication success can then take any desired value if the original
study is significant and the replication sample size is large enough. Compared
to the standard approach to require statistical significance of both the
original and replication study, replication success at the golden level offers
uniform gains in project power and controls the Type I error rate if the
replication sample size is not smaller than the original one. An application to
data from four large replication projects shows that the new approach leads to
more appropriate inferences, as it penalizes shrinkage of the replication
estimate compared to the original one, while ensuring that both effect estimates
are sufficiently convincing on their own.

\textbf{Key words}: Power, replication studies, sceptical $p$-value, shrinkage,
two-trials rule, type I error rate

@@

In this paper we first review the \citet{Held2020} approach for the assessment
of replication success, followed by showing how it relates to the relative
effect size (Section \ref{sec1:res}). This perspective is used in Section
\ref{sec1:goldenthresh} and \ref{sec1:recalib} to propose a recalibration of the
method, the \textit{golden level}, which leads to a more appropriate criterion
for replication success compared to the two-trials rule (Section \ref{sec1:2TR}).
In Section \ref{sec1:ER} we study power and Type I error rates of the proposed
method and compare it to the two-trials rule. The recalibrated method ensures
that conditional power can take any desired value if the original study has been
significant and the replication sample size is large enough (Section
\ref{sec1:powerrep}), controls the overall Type I error if the replication sample
size is not smaller than the original one (Section \ref{sec1:T1E}), and offers
uniform gains in project power compared to the two-trials rule (Section
\ref{sec1:PP}). Section \ref{sec1:application} describes an application to data
from four replication projects and Section \ref{sec1:discussion} closes with some
discussion.

@@

\section{Power and Type I Error Rate}\label{sec1:ER}
Although Bayesian methods do not rely on the frequentist paradigm of repeated
testing, it is still useful to investigate their frequentist operating
characteristics \citep{Dawid1982, Rubin1984, Grieve2016} and this also holds for
the proposed reverse-Bayes assessment of replication success. We first condition
on the results from the original study and compare the power to achieve
replication success with the two-trials rule in Section \ref{sec1:powerrep}. We
then assume that none of the two studies have been conducted and investigate the
overall Type I error rate (Section \ref{sec1:T1E}) and the project power (Section
\ref{sec1:PP}) \citep{Maca2002} over both studies in combination for fixed
relative sample size $c$.

@@

\subsection{Overall Type I error rate}\label{sec1:T1E}
The two studies are assumed to be independent with Type I error rate fixed at
$\alpha$ for each of them, so the Type I error rate of the two-trials rule over
the entire project is simply $\alpha^2$ for any value of the relative effect
size $c$. In contrast, the Type I error rate of the proposed replication success
assessment {depends on} the relative sample size $c$.

For $c=1$, \citet[Section 3]{Held2020} showed that $z_S^2$ in
\eqref{eq1:extrinsic.p2} simplifies to half the harmonic mean of the squared test
statistics $z_o^2$ and $z_r^2$. The connection $z_S^2 = z_H^2/4$ to the harmonic
mean $\chi^2$-test statistic $z_H^2$ \citep{Held2020b}, which has a
$\chi^2(1)$-distribution under the null hypothesis, makes it straightforward to
compute the Type I error rate at level $\alpha_S$ for $c = 1$ as
\begin{eqnarray}\label{eq1:T1E}
\mbox{T1E} =   \left\{1-\Phi\left[2 \, \Phi^{-1}\left(1-\alpha_S \right)
 \right]\right\}/2.
\end{eqnarray}
For the golden level $\alpha_S =0.062$ at $\alpha = 0.025$, the Type I error
rate \eqref{eq1:T1E} is $0.0515$\%, slightly less than the Type I error rate
$\alpha^2=0.0625$\% of the two-trials rule. For comparison, the Type I error
rate at the nominal level $\alpha_S=0.025$ is $0.0022$\%, much smaller than
0.0625\%.

For $c \neq 1$, the Type I error rate can be calculated through numerical
integration:
\begin{equation}\label{eq1:T1E0}
  \mbox{T1E} = \int_{z_{\alpha_S}}^{\infty}
\P(z_r \geq \zrmin \given z_o, c, \alpha_S) \,
  \phi(z_o) \, dz_o,
\end{equation}
where $\phi(\cdot)$ denotes the standard normal density function. The first term
in the integral of \eqref{eq1:T1E0} is the probability of replication success at
level $\alpha_S$ conditional on a fixed original test statistic $z_o$ and a
relative sample size $c$. Now $z_r \sim \Nor(0,1)$ under the null hypothesis, so
this term simplifies to
$\P(z_r \geq \zrmin \given z_o, c, \alpha_S) = 1 - \Phi(\zrmin)$ where $\zrmin$
in \eqref{eq1:eq.z} depends on $z_o$, $c$, and $\alpha_S$.

@@

\caption{Overall Type I error rate (left) and project power (right) for fixed
  relative sample size $c$. Results are given for replication success at the
  nominal and golden level and compared with the two-trials rule (2TR) at
  $\alpha = 0.025$. The dashed darkgrey line is the project power at the golden
  level based on significant original studies ($p_o \leq 0.025$). The power of
  the original study is 90\%.}
\label{fig1:fig4}
\end{figure}

The left plot in Figure \ref{fig1:fig4} displays the Type I error rate for
$\alpha=0.025$ as a function of the relative sample size $c$. It can be seen
that the Type I error of the replication success approach decreases with
increasing relative sample size $c$. This also follows from \eqref{eq1:T1E0}
where $\P(z_r \geq \zrmin \given z_o, c, \alpha_S) = 1 - \Phi(\zrmin)$ decreases
with increasing $c$, because $\zrmin$ increases with increasing $c$, see
equation \eqref{eq1:eq.z}.

The Type I error rate of the nominal level is always below the target
$0.0625$\%. Although the Type I error will eventually attain $\alpha^2$ in the
limit $c \downarrow 0$ \citep[Section 3.4]{Held2020}, the nominal level seems to
be too stringent for realistic values of $c$. The Type I error rate of the
golden level is smaller than $0.0625$\% for $c >0.85$. Appropriate Type I error
control is thus ensured even for replication studies where the sample size is
slightly smaller than in the original study.

@@

\end{center}
\caption{Overall Type I error rate if the replication sample size equal to the
  original study ($c=1$). The two-trials rule (2TR) is compared to replication
  success at the golden and nominal level for different values of $\alpha$.}
\label{fig1:fig5}
\end{figure}

Figure \ref{fig1:fig5} compares for $c=1$ the Type I error rate \eqref{eq1:T1E} of
replication success at the golden and at the nominal level with the two-trials
rule for different values of $\alpha$. The Type I error rate of the two-trials
rule is $\alpha^2$ and the replication success approach at the nominal level
always has a much smaller Type I error rate than $\alpha^2$. At the golden level
the Type I error rate of the replication success approach is much closer to
$\alpha^2$, still slightly smaller if $\alpha < 0.058$. For $\alpha = 0.058$ the
Type I error rate is equal to the Type I error rate $0.058^2 = 0.34\%$ of the
two-trials rule and for $\alpha > 0.058$ the Type I error rate is slightly
larger than $\alpha^2$. The Type I error rate for replication success decreases
with increasing $c$, so as long as the replication sample size is not smaller
than the original sample size, Type I error control at $\alpha^2$ is guaranteed
at the golden level for any one-sided level $\alpha < 0.058$.

@@

We further refined the method by proposing the golden level, a new threshold for
replication success. It guarantees that borderline significant original studies
can only be replicated successfully if the replication effect estimate is larger
than the original one. Compared to the two-trials rule, the golden level offers
uniform gains in project power and controls the Type I error rate at any
one-sided level $\alpha < 0.058$ if the replication sample size
is not smaller than the original one. Empirical evaluation of data from four
replication projects highlights that in most cases the methods are in agreement,
however, for the study pairs where the approaches disagree, the replication
success approach seems to lead to more sensible conclusions. The good
performance has been recently confirmed by a comparison of different replication
success metrics through a simulation study in the presence of publication bias
\citep{Muradchanian2021}.

### 2022-12-09 paper3.tex
one-sided sceptical p

\node[rectangle, draw, rounded corners = 0.2em] (sideCond) [below = 2.5em of nr]
{\begin{tabular}{l}
   \multicolumn{1}{c}{\textbf{Constraints}} \\
   {\scriptsize -- type-I error rate requirements} \\
   {\scriptsize -- cost/availability constraints} \\
   {\scriptsize $\phantom{\rightarrow} \hdots$} \\
 \end{tabular}};

>>>>

\node[rectangle, draw, rounded corners = 0.2em] (sideCond) [below = 2.5em of nr]
{\begin{tabular}{l}
   \multicolumn{1}{c}{\textbf{Constraints}} \\
   {\scriptsize -- type I error rate requirements} \\
   {\scriptsize -- cost/availability constraints} \\
   {\scriptsize $\phantom{\rightarrow} \hdots$} \\
 \end{tabular}};

### 2022-12-09 paper4.tex
one-sided sceptical p

The common 'forward-Bayes' approach is to compute the $\FPR$ from the prior
probability $\Pr(H_0)$ and the Bayes factor with \eqref{eq4:bayes}. However, the
prior probability $\Pr(H_0)$ is usually unknown in practice and often hard to
assess. This can be resolved via the Reverse-Bayes approach
\citep{Colquhoun2017,Colquhoun2019}: Given a $p$-value and a false positive risk
value, calculate the corresponding prior probability $\Pr(H_0)$ that is needed
to achieve that false positive risk. Of specific interest is the value FPR =
5\%, because many scientists believe that a Type-I error of 5\% is equivalent to
a FPR of 5\% \citep{Greenland2016}. This is of course not true and we follow
Example 1 from \citet{Berger1987} and use the Reverse-Bayes approach to derive
the necessary prior assumptions on $\Pr(H_0)$ to achieve FPR = 5\% with Equation
\eqref{eq4:bayes}:
\begin{equation}
  \label{eq4:prior.prob}
     \Pr(H_0) = \left[ 1 + \frac{1-\FPR}{\FPR} \cdot \mbox{BF}_{01}  \right]^{-1}.
\end{equation}

>>>>

The common 'forward-Bayes' approach is to compute the $\FPR$ from the prior
probability $\Pr(H_0)$ and the Bayes factor with \eqref{eq4:bayes}. However, the
prior probability $\Pr(H_0)$ is usually unknown in practice and often hard to
assess. This can be resolved via the Reverse-Bayes approach
\citep{Colquhoun2017,Colquhoun2019}: Given a $p$-value and a false positive risk
value, calculate the corresponding prior probability $\Pr(H_0)$ that is needed
to achieve that false positive risk. Of specific interest is the value FPR =
5\%, because many scientists believe that a Type I error of 5\% is equivalent to
a FPR of 5\% \citep{Greenland2016}. This is of course not true and we follow
Example 1 from \citet{Berger1987} and use the Reverse-Bayes approach to derive
the necessary prior assumptions on $\Pr(H_0)$ to achieve FPR = 5\% with Equation
\eqref{eq4:bayes}:
\begin{equation}
  \label{eq4:prior.prob}
     \Pr(H_0) = \left[ 1 + \frac{1-\FPR}{\FPR} \cdot \mbox{BF}_{01}  \right]^{-1}.
\end{equation}

### 2022-12-09 preface.tex
nicer toc
Lisa feedback
one-sided sceptical p

\section*{Preface}
\addcontentsline{toc}{chapter}{Preface}

This thesis is submitted under the PhD program ``Epidemiology and
Biostatistics'' at the University of Zurich for the degree of Doctor of
Philosophy.
%(\url{http://www.epibiostat-phd.uzh.ch})
The research contained in this thesis was conducted between October 2019 and
December 2022. Financial support was provided by the Swiss National Science
Foundation through the project ``Reverse-Bayes design and analysis of
replication studies'' (project
\href{https://data.snf.ch/grants/grant/189295}{\#189295}) awarded to Leonard
Held.

@@

I also want to thank my friends and colleagues from the Epidemiology,
Biostatistics and Prevention Institute from the University of Zurich (in
alphabetical order): Ainesh, Alexandra, Annina, Charlotte, Babette, Bálint,
Dafne, Dominik, Eveline, Eva, Franscesca, Felix, Julia, Klaus, Kelly, Lucas,
Lisa, Luisa, Goscha, Manuela, Monika, Muriel, Manja, Maria, Marielena, Martin,
Mina, Nadja, Ruedi, Rachel, Sandra, Sona, Steffi, Tala, Torsten, and Ulrike.
Another thanks goes to my friends from the Department of Psychological Methods
from the University of Amsterdam (in alphabetical order): Adam, Alexander,
Alexandra, Alessandra, Alejandro, Angelika, Bruno, Don, František, Frederik,
Jason, Johnny, Joris, Jill, Julia, Lukas, Maarten, Michelle, Nora, Omid,
Quentin, René, Serjan, Suzanne, and Ting. I also want to thank my good friends
from outside of academia (in alphabetical order): Chronis, Dani, Eleftheria,
Flo, Fabi, Giuachin, Mirela, and Peter. A special thanks goes to Ada. Finally, I
thank my family: Beni, Christa, Elisabeth, Fabian, Harry, Markus, Maja, and
Noemi for their support.

>>>>

\section*{Preface}
\addcontentsline{toc}{chapter}{\texorpdfstring{\phantom{IV}Preface}{Preface}}

This thesis is submitted under the PhD program ``Epidemiology and
Biostatistics'' at the University of Zurich for the degree of Doctor of
Philosophy.
%(\url{http://www.epibiostat-phd.uzh.ch})
The research contained in this thesis was conducted between October 2019 and
December 2022. Financial support was provided by the Swiss National Science
Foundation through the project ``Reverse-Bayes design and analysis of
replication studies'' (project
\href{https://data.snf.ch/grants/grant/189295}{\#189295}) awarded to Leonhard
Held.

@@

I also want to thank my friends and colleagues from the Epidemiology,
Biostatistics and Prevention Institute from the University of Zurich (in
alphabetical order): Ainesh, Alexandra, Annina, Charlotte, Babette, Bálint,
Beate, Beni, Dafne, Dominik, Eveline, Eva, Franscesca, Felix, Julia, Klaus,
Kelly, Lucas, Lisa, Luisa, Goscha, Manuela, Monika, Muriel, Manja, Maria,
Marielena, Martin, Mina, Nadja, Ruedi, Rachel, Sandra, Sona, Steffi, Tala,
Torsten, and Ulrike. Another thanks goes to my friends from the Department of
Psychological Methods from the University of Amsterdam (in alphabetical order):
Adam, Alexander, Alexandra, Alessandra, Alejandro, Angelika, Bruno, Don,
František, Frederik, Jason, Johnny, Joris, Jill, Julia, Lukas, Maarten,
Michelle, Nora, Omid, Quentin, René, Serjan, Suzanne, and Ting. I also want to
thank my good friends from outside of academia (in alphabetical order): Chronis,
Dani, Eleftheria, Flo, Fabi, Giuachin, Mirela, and Peter. A special thanks goes
to Ada. Finally, I thank my family: Beni, Christa, Elisabeth, Fabian, Harry,
Markus, Maja, and Noemi for their support.
